---
title: "Interactive ggplots"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#install.packages("animint2")#from CRAN (stable).
##OR from GitHub (development).
#if(!require("remotes"))install.packages("remotes")
#remotes:install_github("tdhock/animint2")
#install.packages("remote")
```


```{r message=FALSE, warning=FALSE}
library(animint2)
library(tidyverse)
library(janitor)
library(here)
library(ggmap)
library(albersusa)
library(skimr)
```

```{r}
data(WorldBank, package="animint2")
tail (WorldBank)
```


```{r}
WorldBank1975 <- subset(WorldBank, year==1975)
head(WorldBank1975)
```


```{r}

scatter1 <- ggplot()+
  geom_point(
    mapping=aes(x=life.expectancy, y=fertility.rate, color=region),
    data=WorldBank1975)
scatter1

#animint(scatter1)
```


```{r}
WorldBankBefore1975 <- subset(WorldBank, 1970 <= year & year <= 1975)
two.layers <- scatter1+
  geom_path(aes(x=life.expectancy, y=fertility.rate, color=region,
                group=country),
            data=WorldBankBefore1975)

two.layers
#(viz.two.layers <- animint(two.layers)) Generates an error: 'unitType' is not an exported object from 'namespace:grid'
```

```{r}
three.layers <- two.layers+
  geom_text(aes(x=life.expectancy, y=fertility.rate, color=region,
                label=country),
            data=WorldBank1975)

three.layers
```

```{r}
two.layersstorage <- two.layers
two.layers$timeSeries <- ggplot()+
  geom_line(aes(x=year, y=fertility.rate, color=region, group=country),
            data=WorldBank)

two.layersstorage
two.layers$timeSeries
```

```{r}
add.x.var <- function(df, x.var){
  data.frame(df, x.var=factor(x.var, c("life expectancy", "year")))
}

  scatterfirst=ggplot()+
    geom_point(aes(x=life.expectancy, y=fertility.rate, color=region),
               data=add.x.var(WorldBank1975, "life expectancy"))+
    geom_path(aes(x=life.expectancy, y=fertility.rate, color=region,
                  group=country),
              data=add.x.var(WorldBankBefore1975, "life expectancy"))+
    geom_line(aes(x=year, y=fertility.rate, color=region, group=country),
              data=add.x.var(WorldBank, "year"))+theme(axis.text.x = element_text(angle = 50, hjust = 1))+
    xlab("")+
    facet_grid(. ~ x.var, scales="free")+
    theme_bw()+
    theme(panel.margin=grid::unit(0, "lines"))
  
  scatterfirst
```


```{r}
show.point.list <- list()
show.path.list <- list()
for(show.year in c(1975, 1985, 1995)){
  show.point.list[[paste(show.year)]] <- data.frame(
    show.year, subset(WorldBank, year==show.year))
  show.path.list[[paste(show.year)]] <- data.frame(
    show.year, subset(WorldBank, show.year - 5 <= year & year <= show.year))
}
show.point <- do.call(rbind, show.point.list)
show.path <- do.call(rbind, show.path.list)
```


```{r}
threepanelsvisual <- list(
  scatter=ggplot()+
    geom_point(aes(x=life.expectancy, y=fertility.rate, color=region),
               data=show.point)+
    geom_path(aes(x=life.expectancy, y=fertility.rate, color=region,
                  group=country),
              data=show.path)+
    facet_grid(. ~ show.year)+
    theme_bw()+
    theme(panel.margin=grid::unit(0, "lines"))
  )

threepanelsvisual
```

## Exercises

Visualize a different type of data not present in the animint2 Manual. 

Data Source: https://rcweb.dartmouth.edu/~f002d69/workshops/index_rspatial.html which represents sightings of grizzly bears (Ursos arctos) in Alaska. 

Other data set used is from Brandell, Ellen E (2021), Serological dataset and R code for: Patterns and processes of pathogen exposure in gray wolves across North America, Dryad, (https://doi.org/10.5061/dryad.5hqbzkh51).  

Load the `grizzly` data and evaluate its structure. 
```{r}
grizzly <- read_csv(here("data", "bear-sightings.csv")) 
grizzly <- janitor::clean_names(grizzly)
grizzly
```

```{r}
glimpse(grizzly)
summary(grizzly)
str(grizzly)
skim(grizzly)
```

Use the range of the latitude and longitude to build an appropriate bounding box for a map.
```{r}
latitude_me <- c(55.02,70.37)
longitude_me <- c(-166.2, -131.3)
bear_box <- make_bbox(longitude_me, latitude_me, f=0.05)
```

Load a map from `stamen` in a terrain style projection and display the map.
```{r}
map_projection <- get_map(bear_box, maptype = "terrain", source = "stamen")
```

```{r}
ggmap(map_projection)
```

Build a final map that overlays the recorded observations of grizzly bears in Alaska.
```{r}
ggmap(map_projection) + geom_point(data = grizzly, aes(longitude,latitude), size=1, color = "blue") + labs(x = "Longitude", y= "Latitude", title = "Recorded Observations of Grizzly Bears in Alaska")
```

Exercise 2: Wolves data. Load the data and evaluate its structure.
```{r}
wolves <- readr::read_csv("data/wolves_data/wolves_dataset.csv")
str(wolves)
skim(wolves)
str(wolves)
```

Evaluating how many distinct wolf populations are included in this study:
```{r}
wolves$pop <- as.factor(wolves$pop)
n_distinct(wolves$pop)
```

```{r}
wolves_US <- wolves %>%
  filter(lat<=48)
wolves_US
```

Use the `albersusa` package to make a base map of the lower 48 US states.
```{r}
us_lower <- usa_sf()
```

```{r}
c_comp <- counties_sf()
```

```{r}
ggplot() +
  geom_sf(data=us_lower, size = 0.25) + theme_linedraw() + labs(title = "US State Boundaries")
```

Use the relimited data to plot the distribution of wolf populations in the lower 48 US states.

```{r}
ggplot() + 
  geom_sf(data = us_lower, size = 0.125) + 
  theme_linedraw()+
  geom_point(data = wolves_US, aes(long, lat, color=pop), shape=2, size=4)+
  labs(title = "Distribution of Wolf Species",
       x = "Longitude",
       y = "Latitude")+
  theme_minimal()
```


Average pack size for the wolves in this study by region:
```{r}
wolves %>%
  group_by(pop) %>%
  summarize(avg_wolfpacksize = mean(pack.size), .groups='keep') %>%
  arrange(desc(avg_wolfpacksize))
```

Make a new map that shows the distribution of wolves in the lower 48 US states but which has the size of location markers adjusted by pack size.
```{r}
wolfplot <- ggplot() +
  geom_sf(data=us_lower, size=0.2) + 
  geom_point(data=wolves_US, aes(long,lat,size=pack.size,color=pack.size), shape=5) + theme_minimal()+
  labs(title = "Distribution of Wolves in Lower 48 US states by Pack Size", x = "Longitude", y="Latitude")

wolfplot
```

```{r}

```



